Quantifying the effects of single mutations on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin

### 2018-01-21 paper/manuscript.tex
Fixed typo in Discussion
revision and response about neutralization curves: show replicates, fit IC50s, test significance

\usepackage{seqsplit}
\usepackage{placeins}

@@

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/avgfracsurvive.pdf}}
\caption{
\label{fig:avgfracsurvive}
{\bf Strain-specific and anti-receptor-binding-site antibodies select mutations with large antigenic effects, but anti-stalk antibodies only select small-effect mutations.}
The excess fraction of virions with a mutation at each site that survive the antibody, averaging across all amino-acid mutations at each site (see Equation~\ref{eq:avgfracsurvive}).
There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1 -- but none for FI6v3 and C179.
Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the largest-effect mutation at each site.
Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show all mutations using logo plots.
Sites are labeled in H3 numbering.
}
\end{figure}

@@

In contrast, the stalk-targeting antibodies C179 and FI6v3 select no strong escape mutants. 
If we look at the results for these antibodies on the same scale as the other antibodies, we see only a few small bumps in the fraction of virions surviving (Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive}).
Only if we zoom in can we see that there are actually a few sites where mutations slightly increase the fraction of virions surviving C179 and FI6v3 (Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo}).
But the effect sizes of these antigenic mutations are tiny compared to the other antibodies -- especially for FI6v3.
Therefore, the HA of A/WSN/1933 influenza virus is far less capable of escaping these anti-stalk antibodies by single mutations than it is of escaping the other four antibodies.

@@

\subsection*{Neutralization assays validate the mutational antigenic profiling}
Do the mutations identified in our mutational antigenic profiling actually have the expected effect on antibody neutralization?
We have previously validated many of the large-effect antigenic mutations selected by the narrow antibodies H17-L19, H17-L10, and H17-L7\cite{doud2017complete}.
However, the mutations selected by the broad anti-stalk antibodies have much smaller effects in our mutational antigenic profiling -- especially for the broadest antibody, FI6v3.
We therefore tested some of these FI6v3-selected mutations using neutralization assays on individual viral mutants.

\begin{figure}
\centerline{\includegraphics[width=0.8\textwidth]{figs/FI6v3mutant_neutcurves/FI6v3_mutant_neutcurves.pdf}}
\caption{
\label{fig:FI6v3neutcurves}
{\bf The mutations selected by FI6v3 increase neutralization resistance, but the effects are small.}
(A) Neutralization curves of individual viral mutants with FI6v3.
The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A). 
The mutations K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization based on the mutational antigenic profiling (Figure~\ref{suppfig:FI6v3logo}).
(B), (C) In contrast to FI6v3, mutations selected by narrow antibodies have very large effects on neutralization.
Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from Doud et al\cite{doud2017complete}.
}
\end{figure}

Figure~\ref{fig:FI6v3neutcurves} shows that the mutational antigenic profiling is highly predictive of the results of the neutralization assays, even for small-effect mutations.
As discussed in the previous section, the three mutations most strongly selected by FI6v3 introduce glycosylation motifs at sites 278-280 or 289-291 (Figure~\ref{fig:structures}A,B).
We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three modestly increased resistance to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A).
As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A).

Our mutational antigenic profiling also identified several non-glycosylation-motif mutations that were selected by FI6v3.
We validated that one of these mutations, G47R in the HA2 chain, increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A) -- although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
We tested one of these mutations, K(-8)T, and it did appear to lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A), although we are not confident that this effect exceeds the noise in our neutralization assays.
As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance within the sensitivity of our assays.

@@

We can examine these possibilities using deep mutational scanning data that measures the tolerance of HA for each possible amino-acid mutation.
Specifically, we have previously selected our A/WSN/1933 virus HA mutant libraries for variants that can replicate in cell culture, and then used deep sequencing to estimate the preference of each site in HA for each possible amino acid\cite{doud2016accurate}.
Figure~\ref{fig:muttolerance} shows these amino-acid preferences for all sites in HA within 4 angstroms of each broad antibody, with the antigenic effects of the mutations overlaid.
Although some HA sites in the antibody footprints strongly prefer a single amino acid, for all antibodies there are also footprint sites that tolerate a fairly wide range of amino acids.
In most cases the mutations selected by the antibodies occur at these mutationally tolerant sites.
However, there are exceptions -- for instance, the H38S mutation selected by C179 is rather disfavored with respect to viral growth, but has a large enough antigenic effect to still be detected in our mutational antigenic profiling.

@@

Another important caveat that our experiments examine \emph{single} amino-acid mutations to the HA from one influenza virus strain.
The protein evolution literature is full of examples of epistatic interactions that enable multiple mutations to access phenotypes not accessible by single mutations\cite{gong2013stability,harms2014historical,starr2017alternative}.
Such epistasis is relevant to HA's evolution.  
For instance, work by Das et al\cite{das2013defining} suggests that the sequential accumulation of mutations can shift the spectrum of available antibody-escape mutations.
Wu et al\cite{wu2017diversity} have used deep mutational scanning to directly demonstrate that rampant epistasis enables HA's receptor-binding pocket to accommodate combinations of individually deleterious mutations, some of which affect sensitivity to antibodies.
Therefore, our work does not imply any absolute limits on the possibilities for antibody escape when evolution is given sufficient time to explore combinations of mutations.
However, single mutations are the most accessible form of genetic variation, and much of influenza virus's natural antigenic drift involves individual mutations that reduce sensitivity to immunodominant antibody specificities\cite{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
Quantifying the antigenic effects of all such mutations therefore provides a relevant measure of ease of viral antibody escape.

A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus\cite{krammer2015advances,corti2017tackling}.
We have used a new approach to quantify the extent to which this is actually true, and shown that neutralization of an H1 virus by broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
Going forward, we suggest that completely mapping viral escape mutations will be a useful complement to more traditional techniques that simply characterize the breadth of anti-viral antibodies against circulating strains.

@@

\subsection*{Neutralization assays}
We performed neutralization assays using influenza viruses that carried GFP in the PB1 segment.
These PB1flank-eGFP were generated in co-cultures of 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells as described previously\cite{bloom2010permissive}, using the standard bi-directional pHW181-PB2, ..., pHW188-NS reverse-genetics plasmids\cite{hoffmann2000dna} for all genes \emph{except} PB1, plus the pHH-PB1flank-eGFP plasmid\cite{bloom2010permissive}.
Each mutant was generated by repeating this process using a version of the pHW184-HA plasmid that had been engineered by site-directed mutagenesis to carry the indicated mutation.
The neutralization assays themselves were performed by using a plate reader to quantify the GFP signal produced by MDCK-SIAT1-CMV-PB1 cells infected by PB1flank-eGFP virus that had been incubated with the indicated antibody concentration as described previously\cite{hooper2013mutant}.
All neutralization curves represent the mean and standard deviation of three measurements. 
\subsection*{H3 sequence numbering}
Unless otherwise indicated, all residues are numbered in the H3 numbering scheme, with the signal peptide in negative numbers, the HA1 subunit as plain numbers, and the HA2 subunit denoted with ``(HA2)''. 
The conversion between sequential numbering of the A/WSN/1933 HA and the H3 numbering scheme was performed using the Python script available at \url{https://github.com/jbloomlab/HA_numbering}.
File~\ref{suppfile:HAnumbering} gives the numbering conversion.

\subsection*{Inference of HA phylogenetic tree}
To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database\cite{zhang2017influenza}, inferred the phylogenetic tree using RaxML\cite{stamatakis2014raxml} with a GTR model, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}). 
The HA sequences used are in File~\ref{suppfile:HAsubtypes}. 
In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize\cite{yoshida2009cross, lee2012heterosubtypic, okuno1993common, dreyfus2013structure, corti2011neutralizing}. 
Among broad antibodies, S139/1 has not been tested against H8 and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes -- however, since these antibodies have a very limited range even among H1 HAs\cite{caton1982antigenic}, we assume that they do not bind other subtypes.

@@

\subsection*{Analysis of deep sequencing data}
The deep sequencing data were analyzed using version 2.1.0 of the \texttt{dms\_tools2} software package\cite{bloom2015software}, which is available at \url{http://jbloomlab.github.io/dms_tools2}.
File~\ref{suppfile:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
Detailed statistics about the sequencing depth and error rates are shown in this Jupyter notebook and its HTML rendering in File~\ref{suppfile:analysis_html}.

@@

In this paper, Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive} show the median of excess fraction surviving taken across all biological and technical replicates at a given antibody concentration.
The subsequent logo plots show the medians of these values taken across all concentrations for each antibody.
The numerical values plotted in these logo plots are in File~\ref{suppfile:fracsurvive}.

@@

@@

\begin{suppfile}
\caption{\label{suppfile:fracsurvive}
{\bf The excess fraction surviving for each mutation for each antibody.}
This file is a ZIP of CSV files giving the numerical values plotted in the logo plots.
These are median excess fraction surviving taken first across replicates and then across antibody concentrations.
}
\end{suppfile}

>>>>

\usepackage{seqsplit}
\usepackage{placeins}
\usepackage{booktabs}

@@

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/avgfracsurvive.pdf}}
\caption{
\label{fig:avgfracsurvive}
{\bf Strain-specific and anti-receptor-binding-site antibodies select mutations with large antigenic effects, but anti-stalk antibodies only select small-effect mutations.}
The excess fraction of virions with a mutation at each site that survive the antibody, averaging across all amino-acid mutations at each site (see Equation~\ref{eq:avgfracsurvive}).
There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1---but none for FI6v3 and C179.
Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the largest-effect mutation at each site.
Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show all mutations using logo plots.
Sites are labeled in H3 numbering.
}
\end{figure}

@@

In contrast, the stalk-targeting antibodies C179 and FI6v3 select no strong escape mutants. 
If we look at the results for these antibodies on the same scale as the other antibodies, we see only a few small bumps in the fraction of virions surviving (Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive}).
Only if we zoom in can we see that there are actually a few sites where mutations slightly increase the fraction of virions surviving C179 and FI6v3 (Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo}).
But the effect sizes of these antigenic mutations are tiny compared to the other antibodies---especially for FI6v3.
Therefore, the HA of A/WSN/1933 influenza virus is far less capable of escaping these anti-stalk antibodies by single mutations than it is of escaping the other four antibodies.

@@

\subsection*{Neutralization assays validate the mutational antigenic profiling}
Do the mutations identified in our mutational antigenic profiling actually have the expected effect on antibody neutralization?
We have previously validated many of the large-effect antigenic mutations selected by the narrow antibodies H17-L19, H17-L10, and H17-L7\cite{doud2017complete}.
However, the mutations selected by the broad anti-stalk antibodies have much smaller effects in our mutational antigenic profiling---especially for the broadest antibody, FI6v3.
We therefore tested some of these FI6v3-selected mutations using neutralization assays on individual viral mutants.

\begin{figure}
\centerline{\includegraphics[width=0.8\textwidth]{figs/FI6v3mutant_neutcurves/FI6v3_mutant_neutcurves.pdf}}
\caption{
\label{fig:FI6v3neutcurves}
{\bf The mutations selected by FI6v3 increase neutralization resistance, but the effects are small.}
(A) Neutralization curves of individual viral mutants with FI6v3.
The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A), whereas K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization (Figure~\ref{suppfig:FI6v3logo}).
All neutralization curves in this panel were performed in triplicate on the same day.
This panel shows the average of the replicates; Figure~\ref{suppfig:FI6v3replicates} shows the curves for each replicate individually and performs statistical testing of whether the IC50s for mutants are significantly different than for wildtype.
(B), (C) In contrast to FI6v3, mutations selected by narrow antibodies have very large effects on neutralization.
Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from Doud et al\cite{doud2017complete}.
}
\end{figure}

Figure~\ref{fig:FI6v3neutcurves} shows that the mutational antigenic profiling is highly predictive of the results of the neutralization assays, even for small-effect mutations.
As discussed in the previous section, the three mutations most strongly selected by FI6v3 introduce glycosylation motifs at sites 278-280 or 289-291 (Figure~\ref{fig:structures}A,B).
We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three modestly but significantly increased resistance to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates}).
As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not significantly shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates}).

Our mutational antigenic profiling also identified several non-glycosylation-motif mutations that were selected by FI6v3.
We validated that one of these mutations, G47R in the HA2 chain, significantly increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates})---although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
We tested one of these mutations, K(-8)T, and it did lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates})---although despite the significance testing in Figure~\ref{suppfig:FI6v3replicates}, we remain circumspect about the magnitude of this effect relative to the noise in our neutralization assays.
As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them significantly affected neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates}).

@@

We can examine these possibilities using deep mutational scanning data that measures the tolerance of HA for each possible amino-acid mutation.
Specifically, we have previously selected our A/WSN/1933 virus HA mutant libraries for variants that can replicate in cell culture, and then used deep sequencing to estimate the preference of each site in HA for each possible amino acid\cite{doud2016accurate}.
Figure~\ref{fig:muttolerance} shows these amino-acid preferences for all sites in HA within 4 angstroms of each broad antibody, with the antigenic effects of the mutations overlaid.
Although some HA sites in the antibody footprints strongly prefer a single amino acid, for all antibodies there are also footprint sites that tolerate a fairly wide range of amino acids.
In most cases the mutations selected by the antibodies occur at these mutationally tolerant sites.
However, there are exceptions---for instance, the H38S mutation selected by C179 is rather disfavored with respect to viral growth, but has a large enough antigenic effect to still be detected in our mutational antigenic profiling.

@@

Another important caveat is that our experiments examine \emph{single} amino-acid mutations to the HA from one influenza virus strain.
The protein evolution literature is full of examples of epistatic interactions that enable multiple mutations to access phenotypes not accessible by single mutations\cite{gong2013stability,harms2014historical,starr2017alternative}.
Such epistasis is relevant to HA's evolution.  
For instance, work by Das et al\cite{das2013defining} suggests that the sequential accumulation of mutations can shift the spectrum of available antibody-escape mutations.
Wu et al\cite{wu2017diversity} have used deep mutational scanning to directly demonstrate that rampant epistasis enables HA's receptor-binding pocket to accommodate combinations of individually deleterious mutations, some of which affect sensitivity to antibodies.
Therefore, our work does not imply any absolute limits on the possibilities for antibody escape when evolution is given sufficient time to explore combinations of mutations.
However, single mutations are the most accessible form of genetic variation, and much of influenza virus's natural antigenic drift involves individual mutations that reduce sensitivity to immunodominant antibody specificities\cite{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
Quantifying the antigenic effects of all such mutations therefore provides a relevant measure of ease of viral antibody escape.

A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus\cite{krammer2015advances,corti2017tackling}.
We have used a new approach to quantify the extent to which this is actually true, and shown that neutralization of an H1 virus by broad anti-stalk antibodies is indeed more---although certainly not completely---resistant to erosion by viral point mutations.
Going forward, we suggest that completely mapping viral escape mutations will be a useful complement to more traditional techniques that simply characterize the breadth of anti-viral antibodies against circulating strains.

@@

\subsection*{Neutralization assays}
We performed neutralization assays using influenza viruses that carried GFP in the PB1 segment.
These PB1flank-eGFP were generated in co-cultures of 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells as described previously\cite{bloom2010permissive}, using the standard bi-directional pHW181-PB2, ..., pHW188-NS reverse-genetics plasmids\cite{hoffmann2000dna} for all genes \emph{except} PB1, plus the pHH-PB1flank-eGFP plasmid\cite{bloom2010permissive}.
Each mutant was generated by repeating this process using a version of the pHW184-HA plasmid that had been engineered by site-directed mutagenesis to carry the indicated mutation.
The neutralization assays themselves were performed by using a plate reader to quantify the GFP signal produced by MDCK-SIAT1-CMV-PB1 cells infected by PB1flank-eGFP virus that had been incubated with the indicated antibody concentration as described previously\cite{hooper2013mutant}.
All neutralization curves in Figure~\ref{fig:FI6v3neutcurves}A represent the mean and standard deviation of three measurements, with the individual replicates shown in Figure~\ref{suppfig:FI6v3replicates}.
All the neutralization assays for FI6v3 were performed on the same day to eliminate batch effects, with each replicate involving independent serial dilution of the antibody in a separate column of a 96-well plate.

\subsection*{H3 sequence numbering}
Unless otherwise indicated, all residues are numbered in the H3 numbering scheme, with the signal peptide in negative numbers, the HA1 subunit as plain numbers, and the HA2 subunit denoted with ``(HA2)''. 
The conversion between sequential numbering of the A/WSN/1933 HA and the H3 numbering scheme was performed using the Python script available at \url{https://github.com/jbloomlab/HA_numbering}.
File~\ref{suppfile:HAnumbering} gives the numbering conversion.

\subsection*{Inference of HA phylogenetic tree}
To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database\cite{zhang2017influenza}, inferred the phylogenetic tree using RaxML\cite{stamatakis2014raxml} with a GTR model, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}). 
The HA sequences used are in File~\ref{suppfile:HAsubtypes}. 
In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize\cite{yoshida2009cross, lee2012heterosubtypic, okuno1993common, dreyfus2013structure, corti2011neutralizing}. 
Among broad antibodies, S139/1 has not been tested against H8 and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes---however, since these antibodies have a very limited range even among H1 HAs\cite{caton1982antigenic}, we assume that they do not bind other subtypes.

@@

\subsection*{Analysis of deep sequencing data}
The deep sequencing data were analyzed using version 2.2.1 of the \texttt{dms\_tools2} software package\cite{bloom2015software}, which is available at \url{http://jbloomlab.github.io/dms_tools2}.
File~\ref{suppfile:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
Detailed statistics about the sequencing depth and error rates are shown in this Jupyter notebook and its HTML rendering in File~\ref{suppfile:analysis_html}.

@@

In this paper, Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive} show the median of excess fraction surviving taken across all biological and technical replicates at a given antibody concentration (Equation~\ref{eq:fracsurvive_excess}).
The subsequent logo plots show the medians of these values taken across all concentrations for each antibody.
The numerical values plotted in these logo plots are in File~\ref{suppfile:fracsurvive_excess}.
The fraction surviving values \emph{not} adjusted to be in excess of the library average (Equation~\ref{eq:fracsurvive}) are in File~\ref{suppfile:fracsurvive}.

@@

\begin{suppfigure}
{\bf \LARGE A} \\
\centerline{\includegraphics[width=\textwidth]{figs/FI6v3_replicate_neutcurves.pdf}} \\
\vspace{0.1in}
{\bf \LARGE B}
\begin{center}
\input{figs/FI6v3_replicate_IC50s}
\end{center}
\caption{\label{suppfig:FI6v3replicates}
{\bf Replicates of the FI6v3 neutralization curves in Figure~\ref{fig:FI6v3neutcurves}A.}
The neutralization assays were performed in triplicate for all nine mutants and wildtype.
Figure~\ref{fig:FI6v3neutcurves}A shows the \emph{average} of those replicates.
(A) The neutralization data for each replicate shown individually, with IC50 values fit using a four-parameter logistic curve with the top value constrained to one (see \url{https://jbloomlab.github.io/dms_tools2/dms_tools2.neutcurve.html} for the code used for the fitting.)
(B) Table of the IC50 values for each replicate.
We used an unpaired Student's t-test with unequal variances to test the null hypothesis that each mutant had an IC50 indistinguishable from wildtype.
We then used Bonferroni's method to correct the $P$-values for multiple testing, and report these corrected values.
}
\end{suppfigure}

@@

\begin{suppfile}
\caption{\label{suppfile:fracsurvive_excess}
{\bf The excess fraction surviving for each mutation for each antibody.}
This file is a ZIP of CSV files giving the numerical values plotted in the logo plots.
These are median excess fraction surviving taken first across replicates and then across antibody concentrations.
See Equation~\ref{eq:fracsurvive_excess}.
}
\end{suppfile}

\begin{suppfile}
\caption{\label{suppfile:fracsurvive}
{\bf The fraction surviving for each mutation for each antibody.}
This file differs from File~\ref{suppfile:fracsurvive_excess} only in that the values are \emph{not} adjusted to be in excess of the library average (e.g., they are from Equation~\ref{eq:fracsurvive} rather than Equation~\ref{eq:fracsurvive_excess}).
}
\end{suppfile}

### 2018-01-21 paper/manuscript_NatureMicrobiology_submission.tex
revision and response about neutralization curves: show replicates, fit IC50s, test significance

\subsection*{Neutralization assays validate the mutational antigenic profiling}
Do the mutations identified in our mutational antigenic profiling actually have the expected effect on antibody neutralization?
We have previously validated many of the large-effect antigenic mutations selected by the narrow antibodies H17-L19, H17-L10, and H17-L7\cite{doud2017complete}.
However, the mutations selected by the broad anti-stalk antibodies have much smaller effects in our mutational antigenic profiling -- especially for the broadest antibody, FI6v3.
We therefore tested some of these FI6v3-selected mutations using neutralization assays on individual viral mutants.

@@

Our mutational antigenic profiling also identified several non-glycosylation-motif mutations that were selected by FI6v3.
We validated that one of these mutations, G47R in the HA2 chain, increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A) -- although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
We tested one of these mutations, K(-8)T, and it did appear to lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A), although we are not confident that this effect exceeds the noise in our neutralization assays.
As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance within the sensitivity of our assays.

@@

A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus\cite{krammer2015advances,corti2017tackling}.
We have used a new approach to quantify the extent to which this is actually true, and shown that neutralization of an H1 virus by broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
Going forward, we suggest that completely mapping viral escape mutations will be a useful complement to more traditional techniques that simply characterize the breadth of anti-viral antibodies against circulating strains.

@@

\subsection*{Inference of HA phylogenetic tree}
To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database\cite{zhang2017influenza}, inferred the phylogenetic tree using RaxML\cite{stamatakis2014raxml} with a GTR model, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}). 
The HA sequences used are in File~\ref{suppfile:HAsubtypes}. 
In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize\cite{yoshida2009cross, lee2012heterosubtypic, okuno1993common, dreyfus2013structure, corti2011neutralizing}. 
Among broad antibodies, S139/1 has not been tested against H8 and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes -- however, since these antibodies have a very limited range even among H1 HAs\cite{caton1982antigenic}, we assume that they do not bind other subtypes.

>>>>

\subsection*{Neutralization assays validate the mutational antigenic profiling}
Do the mutations identified in our mutational antigenic profiling actually have the expected effect on antibody neutralization?
We have previously validated many of the large-effect antigenic mutations selected by the narrow antibodies H17-L19, H17-L10, and H17-L7\cite{doud2017complete}.
However, the mutations selected by the broad anti-stalk antibodies have much smaller effects in our mutational antigenic profiling---especially for the broadest antibody, FI6v3.
We therefore tested some of these FI6v3-selected mutations using neutralization assays on individual viral mutants.

@@

Our mutational antigenic profiling also identified several non-glycosylation-motif mutations that were selected by FI6v3.
We validated that one of these mutations, G47R in the HA2 chain, increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A)---although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
We tested one of these mutations, K(-8)T, and it did appear to lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A), although we are not confident that this effect exceeds the noise in our neutralization assays.
As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance within the sensitivity of our assays.

@@

A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus\cite{krammer2015advances,corti2017tackling}.
We have used a new approach to quantify the extent to which this is actually true, and shown that neutralization of an H1 virus by broad anti-stalk antibodies is indeed more---although certainly not completely---resistant to erosion by viral point mutations.
Going forward, we suggest that completely mapping viral escape mutations will be a useful complement to more traditional techniques that simply characterize the breadth of anti-viral antibodies against circulating strains.

@@

\subsection*{Inference of HA phylogenetic tree}
To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database\cite{zhang2017influenza}, inferred the phylogenetic tree using RaxML\cite{stamatakis2014raxml} with a GTR model, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}). 
The HA sequences used are in File~\ref{suppfile:HAsubtypes}. 
In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize\cite{yoshida2009cross, lee2012heterosubtypic, okuno1993common, dreyfus2013structure, corti2011neutralizing}. 
Among broad antibodies, S139/1 has not been tested against H8 and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes---however, since these antibodies have a very limited range even among H1 HAs\cite{caton1982antigenic}, we assume that they do not bind other subtypes.
